---
title: "SPL_DataSet_Combine"
author: "Elissa Ewert"
date: "February 9, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=FALSE}
# load packages
library(tidyverse)
library(magrittr)
```

```{r}
# data location

folder <- "C:/Users/Elissa/Documents/DSMS/Capstone_SPL/Data/RawData/"

outFolder <- "C:/Users/Elissa/Documents/DSMS/Capstone_SPL/Data/"
```

# Read in data from SPL

```{r}
# read in OverDrive counts

#read in first data set
overdrive1 <- read.csv(file = paste0(folder, "data_req_overdrive_2018_01_2021_07.csv"), sep = ",")

# convert date to character
overdrive1$date <- as.character(as.Date(overdrive1$date))

# rename column for binding dfs
colnames(overdrive1)[1] <- "branch"

# read in second data set
overdrive2 <- read.csv(file = paste0(folder, "data_req_overdrive_2021_08_2021_12.csv"), sep = "\t")

# convert date format and convetr to character
overdrive2$date <- as.character(as.Date(overdrive2$date, format = "%m/%d/%Y"))

# bind rows and remove dfs
overDrive <- bind_rows(overdrive1, overdrive2)

rm(overdrive1, overdrive2)


# sum checkout and borrower_count data by date
uniqueDate <- unique(overDrive$date)

overDrive$totOverDRiveCko <- NA
overDrive$totOverDRiveBorwCnt <-NA

for(x in uniqueDate){
  overDrive$totOverDRiveCko[overDrive$date == x] = sum(overDrive$checkouts[overDrive$date == x])
  overDrive$totOverDRiveBorwCnt[overDrive$date == x] = sum(overDrive$borrower_count[overDrive$date == x])
}

overDrive %<>% select(c(date, totOverDRiveCko, totOverDRiveBorwCnt))

overDrive <- overDrive[!duplicated(overDrive[,c("date", "totOverDRiveCko", "totOverDRiveBorwCnt")]), ]

row.names(overDrive) <- 1:nrow(overDrive)
```



```{r}
# read in openhours use counts

openhours1 <- read.csv(file = paste0(folder, "data_req_open_hours_2018_01_2021_07.csv"), sep = ",")

openhours1$date <- as.character(as.Date(openhours1$date))

colnames(openhours1)[1] <- "branch"

openhours2 <- read.csv(file = paste0(folder, "data_req_open_hours_2021_08_2021_12.csv"), sep = "\t")

# convert date format and conver to character
openhours2$date <- as.character(as.Date(openhours2$date, format = "%m/%d/%Y"))

# bind rows and remove dfs
openHours <- bind_rows(openhours1, openhours2)

rm(openhours1, openhours2)

# keep only true branches, not dropbox or ILL
keep <- c("BAL", "BEA", "BRO", "CAP", "CEN", "COL", "DLR", "DTH", "FRE", "GLK", "GWD", "HIP", 
          "IDC", "LCY", "MAG", "MGM", "MON", "NET", "NGA", "NHY", "QNA", "RBE", "SPA", "SWT", 
          "UNI", "WAL", "WTS")

openHours <- openHours[openHours$branch %in% keep, ]

rownames(openHours) <- 1:nrow(openHours)

# sum open hours traffic counts data by date
uniqueDate <- unique(openHours$date)
uniqueBranch <- unique(openHours$branch)

openHours$totHours <- NA
openHours$totBranchOpen <- NA
openHours$pubHealthClosure <- NA
openHours$closedNotes <- NA

for(x in uniqueDate){
  openHours$totHours[openHours$date == x] = sum(openHours$open_hours_count[openHours$date == x])
  branchesOpen <- 0
  pubHealth <-0
  closedNote = ""
  for(z in uniqueBranch){
    if(openHours$open_hours_count[openHours$branch == z & openHours$date == x] > 0){
      branchesOpen = branchesOpen +1
    }
    if(openHours$closed_note[openHours$branch == z & openHours$date == x] == "Public Health Closure"){
      pubHealth = pubHealth + 1
    }  
    if(openHours$closed_note[openHours$branch == z & openHours$date == x] != "NULL"){
      closedNote = paste(closedNote, openHours$closed_note[openHours$branch == z & openHours$date == x])
    }
  }
  openHours$closedNotes[openHours$date == x] <- closedNote 
  openHours$totBranchOpen[openHours$date == x] <- branchesOpen
  openHours$pubHealthClosure[openHours$date == x] <- pubHealth
}

openHours %<>% select(c(date, totHours, totBranchOpen, pubHealthClosure, closedNotes))

openHours <- openHours[!duplicated(openHours[,c("date", "totHours", "totBranchOpen", "pubHealthClosure", "closedNotes")]), ]

row.names(openHours) <- 1:nrow(openHours)

```


```{r}
# read in computer use counts

#read in first data set
computeruse1 <- read.csv(file = paste0(folder, "data_req_computer_use_2018_01_2021_07.csv"), sep = ",")

#convert date to character
computeruse1$date <- as.character(as.Date(computeruse1$date))

#fix column name for binding dfs
colnames(computeruse1)[1] <- "branch"

#read in second data set
computeruse2 <- read.csv(file = paste0(folder, "data_req_computer_use_2021_08_2021_12.csv"), sep = "\t")

# convert date format and conver to character
computeruse2$date <- as.character(as.Date(computeruse2$date, format = "%m/%d/%Y"))

# bind rows and remove dfs
computerUse <- bind_rows(computeruse1, computeruse2)

rm(computeruse1, computeruse2)


# sum pc_session data by date
uniqueDate <- unique(computerUse$date)

computerUse$totPCsessions <- NA

for(x in uniqueDate){
  computerUse$totPCsessions[computerUse$date == x] = sum(computerUse$pc_sessions[computerUse$date == x])
}

computerUse = merge(x = computerUse, y = openHours, by="date", all.x = TRUE)

computerUse %<>% select(c(date, totPCsessions))

computerUse <- computerUse[!duplicated(computerUse[,c("date", "totPCsessions")]), ]

row.names(computerUse) <- 1:nrow(computerUse)
```



```{r}
# read in in-person programs use counts - No in-person programs in second date range

inPersonPrograms <- read.csv(file = paste0(folder, "data_req_inperson_programs_2018_01_2021_07.csv"), sep = ",")

inPersonPrograms$date <- as.character(as.Date(inPersonPrograms$date))

colnames(inPersonPrograms)[1] <- "branch"


# sum program and attendance data by month
uniqueDate <- unique(inPersonPrograms$date)

inPersonPrograms$totProg <- NA
inPersonPrograms$totAttnd <-NA

inPersonPrograms$attendance <- as.numeric(inPersonPrograms$attendance)

for(x in uniqueDate){
  inPersonPrograms$totProg[inPersonPrograms$date == x] = sum(inPersonPrograms$program_count[inPersonPrograms$date == x])
  inPersonPrograms$totAttnd[inPersonPrograms$date == x] = sum(inPersonPrograms$attendance[inPersonPrograms$date == x])
}

inPersonPrograms %<>% select(c(date, month, year, totProg, totAttnd))

inPersonPrograms <- inPersonPrograms[!duplicated(inPersonPrograms[,c("date", "totProg", "totAttnd")]), ]

row.names(inPersonPrograms) <- 1:nrow(inPersonPrograms)

```


```{r}
# read in traffic counts

traffic1 <- read.csv(file = paste0(folder, "data_req_traffic_2018_01_2021_07.csv"), sep = ",")

traffic1$date <- as.character(as.Date(traffic1$date))

colnames(traffic1)[1] <- "branch"

traffic2 <- read.csv(file = paste0(folder, "data_req_traffic_2021_08_2021_12.csv"), sep = "\t")

# convert date format and conver to character
traffic2$date <- as.character(as.Date(traffic2$date, format = "%m/%d/%Y"))

# bind rows and remove dfs
traffic <- bind_rows(traffic1, traffic2)

rm(traffic1, traffic2)


# sum open hours traffic counts data by month
uniqueDate <- unique(traffic$date)

traffic$totTraffic <- NA

for(x in uniqueDate){
  traffic$totTraffic[traffic$date == x] = sum(traffic$open_hours_traffic_count[traffic$date == x])
}

traffic %<>% select(c(date, totTraffic))

traffic <- traffic[!duplicated(traffic[,c("date", "totTraffic")]), ]

row.names(traffic) <- 1:nrow(traffic)

```


```{r}
# read in wifi counts

wifi1 <- read.csv(file = paste0(folder, "data_req_wifi_2018_01_2021_07.csv"), sep = ",")

# date as character
wifi1$date <- as.character(as.Date(wifi1$date))

#fix column name for binding data frames
colnames(wifi1)[1] <- "branch"

# read in second data set
wifi2 <- read.csv(file = paste0(folder, "data_req_wifi_2021_08_2021_12.csv"), sep = "\t")

# convert date format and conver to character
wifi2$date <- as.character(as.Date(wifi2$date, format = "%m/%d/%Y"))

# bind rows and remove dfs
wifi <- bind_rows(wifi1, wifi2)

rm(wifi1, wifi2)


# sum connections and MB used data by day
uniqueDate <- unique(wifi$date)

wifi$totconnect <- NA
wifi$totMBused <-NA

for(x in uniqueDate){
  wifi$totconnect[wifi$date == x] = sum(wifi$connections[wifi$date == x])
  wifi$totMBused[wifi$date == x] = sum(wifi$mb_used[wifi$date == x])
}

wifi %<>% select(c(date, totconnect, totMBused))

wifi <- wifi[!duplicated(wifi[,c("date", "totconnect", "totMBused")]), ]

row.names(wifi) <- 1:nrow(wifi)

```

# Read in circulation data from Seattle Open Data - extracted from original file using Python script

```{r}
# circulation data from Seattle Open Data 

printCirc <- read.csv(file = paste0(outFolder, "allCircDOW_18-21_df.csv"), sep = ",", )

printCirc %<>% select(-c(X))

printCirc[is.na(printCirc)] <- 0

printCirc$otherType <- NA

#compbine small value item type  into "other" column
for(x in 1:nrow(printCirc)){
  printCirc$otherType[x] <- sum(printCirc[x, c(4,8, 12:42)])
}

printCirc %<>% select(c(date, totCirc, acbk, jcbk, accd, pkbknh, acdvd, jcdvd, jccd, otherType, nameOfDay))

printCirc$date <- as.character(as.Date(printCirc$date, format = "%m/%d/%Y"))
```

# join library data into single data set

```{r}
data = merge(x = printCirc, y = openHours, by="date", all.x = TRUE)
data = merge(x = data, y = wifi, by="date", all.x = TRUE)
data = merge(x = data, y = traffic, by="date", all.x = TRUE)
data = merge(x = data, y = inPersonPrograms, by="date", all.x = TRUE) #monthly statistic
data = merge(x = data, y = computerUse, by="date", all.x = TRUE)
data = merge(x = data, y = overDrive, by="date", all.x = TRUE)

rm(printCirc, openHours, wifi, traffic, inPersonPrograms, computerUse, overDrive)

data <- data[order(data$date), ]
```

# Read in Covid19 data from King County Public Health and join with SPL data

```{r}
# read in covid19 counts

covid19 <- read.csv(file = paste0(folder, "king_county_Covid_data.csv"), sep = ",")

colnames(covid19)[1] <- "City"
colnames(covid19)[2] <- "date"

# convert date format and conver to character
covid19$date <- as.character(as.Date(covid19$date, format = "%m/%d/%Y"))


#sum data by data for full county
uniqueDate <- unique(covid19$date)

covid19$allTestResultsCnty <- NA
covid19$peopleTestedCnty <-NA
covid19$positivesCnty <-NA
covid19$confirmedCasesCnty <-NA
covid19$probableCasesCnty <-NA
covid19$hospitalizationsCnty <-NA
covid19$deathsCnty <-NA

for(x in uniqueDate){
  covid19$allTestResultsCnty[covid19$date == x] = sum(covid19$All_Test_Results[covid19$date == x])
  
  covid19$peopleTestedCnty[covid19$date == x] = sum(covid19$People_Tested[covid19$date == x])
  
  covid19$positivesCnty[covid19$date == x] = sum(covid19$Positives[covid19$date == x])
  
  covid19$confirmedCasesCnty[covid19$date == x] = sum(covid19$Confirmed_Cases[covid19$date == x])
  
  covid19$probableCasesCnty[covid19$date == x] = sum(covid19$Probable_Cases[covid19$date == x])
  
  covid19$hospitalizationsCnty[covid19$date == x] = sum(covid19$Hospitalizations[covid19$date == x])
  
  covid19$deathsCnty[covid19$date == x] = sum(covid19$Deaths[covid19$date == x])
}

covid19cnty <- select(covid19, c(date, allTestResultsCnty, peopleTestedCnty, positivesCnty, confirmedCasesCnty,          
                          probableCasesCnty, hospitalizationsCnty, deathsCnty ))

covid19cnty <- covid19cnty[!duplicated(covid19cnty[,c('date', "allTestResultsCnty", "peopleTestedCnty", "positivesCnty", 
                                              'confirmedCasesCnty', "probableCasesCnty", "hospitalizationsCnty", 
                                              "deathsCnty" )]), ]

row.names(covid19cnty) <- 1:nrow(covid19cnty)


#sum data by data for Seattle only
uniqueDate <- unique(covid19$date)

covid19$allTestResultsStl <- NA
covid19$peopleTestedStl <-NA
covid19$positivesStl <-NA
covid19$confirmedCasesStl <-NA
covid19$probableCasesStl <-NA
covid19$hospitalizationsStl <-NA
covid19$deathsStl <-NA

for(x in uniqueDate){
  covid19$allTestResultsStl[covid19$date == x] = sum(covid19$All_Test_Results[covid19$date == x & covid19$City == "Seattle"])
  
  covid19$peopleTestedStl[covid19$date == x] = sum(covid19$People_Tested[covid19$date == x & covid19$City == "Seattle"])
  
  covid19$positivesStl[covid19$date == x] = sum(covid19$Positives[covid19$date == x & covid19$City == "Seattle"])
  
  covid19$confirmedCasesStl[covid19$date == x] = sum(covid19$Confirmed_Cases[covid19$date == x & covid19$City == "Seattle"])
  
  covid19$probableCasesStl[covid19$date == x] = sum(covid19$Probable_Cases[covid19$date == x & covid19$City == "Seattle"])
  
  covid19$hospitalizationsStl[covid19$date == x] = sum(covid19$Hospitalizations[covid19$date == x & covid19$City == "Seattle"])
  
  covid19$deathsStl[covid19$date == x] = sum(covid19$Deaths[covid19$date == x & covid19$City == "Seattle"])
}

covid19stl <- select(covid19, c(date, allTestResultsStl, peopleTestedStl, positivesStl, confirmedCasesStl,          
                          probableCasesStl, hospitalizationsStl, deathsStl ))

covid19stl <- covid19stl[!duplicated(covid19stl[,c('date', "allTestResultsStl", "peopleTestedStl", "positivesStl", 
                                              'confirmedCasesStl', "probableCasesStl", "hospitalizationsStl", 
                                              "deathsStl" )]), ]

row.names(covid19stl) <- 1:nrow(covid19stl)
```


```{r}
data = merge(x = data, y = covid19stl, by="date", all.x = TRUE)
data = merge(x = data, y = covid19cnty, by="date", all.x = TRUE)

covidColumns <- c("allTestResultsStl", "peopleTestedStl", "positivesStl", 'confirmedCasesStl', "probableCasesStl",
                  "hospitalizationsStl", "deathsStl", "allTestResultsCnty", "peopleTestedCnty", "positivesCnty",
                  'confirmedCasesCnty', "probableCasesCnty", "hospitalizationsCnty", "deathsCnty")

# replace NAs for days before Feb 2020 with 0s (pre-covid, not numbers)
for(c in covidColumns){
  is.na(data[, c]) <- 0
}

data <- data[order(data$date), ]
```

```{r}
#read in vaccine data
vaccine <- read.csv(file = paste0(folder, "KingCounty_Vaccine_Day.csv"), sep = ",")

vaccine$date <- as.character(vaccine$date)
vaccine$mRNA_1 <- vaccine$Primary.mRNA.dose.1
vaccine$mRNA_2 <- vaccine$Primary.mRNA.dose.2
vaccine$JJdose <- vaccine$Primary.J.J.Janssen.dose
vaccine$booster <- vaccine$Additional.booster.dose
vaccine %<>% select(c(date, mRNA_1, mRNA_2, booster, JJdose, Population))

data = merge(x = data, y = vaccine, by="date", all.x = TRUE)
```

# Add data for Stay at Home order for WA and status of pandemic (Y/N) from WHO

```{r}
# create column for day number of pandemic and 

date <- seq(as.Date("2020-03-11"), as.Date("2021-12-31"), by="days")

pandemicDayNumber <- seq(1,length(date), by = 1)

pandemicYes <- rep("Y",length(date))

worldWidePandemic <- cbind(data.frame(date), data.frame(pandemicDayNumber), data.frame(pandemicYes))

worldWidePandemic$date %<>% as.character

# column for stay at home order
date <- seq(as.Date("2020-03-25"), as.Date("2020-06-30"), by="days")

stayAtHome <- rep("Y",length(date))

WAstayAtHome <- cbind(data.frame(date), data.frame(stayAtHome))

WAstayAtHome$date %<>% as.character

#merge
data = merge(x = data, y = worldWidePandemic, by="date", all.x = TRUE)

data$pandemicYes[is.na(data$pandemicYes)] <- "N"

data = merge(x = data, y = WAstayAtHome, by="date", all.x = TRUE)

data$stayAtHome[is.na(data$stayAtHome)] <- "N"

rm(covid19cnty, covid19stl, vaccine, worldWidePandemic, WAstayAtHome, covid19)
```


#recode column names for clarity

```{r}
colnames(data) %<>% recode("totCirc" = "circ_total_physical",
                           "acbk" = "circ_acbk",
                           "jcbk" = "circ_jcbk",
                           "pkbknh" = "circ_pkbknh",
                           "acdvd" = "circ_acdvd",
                           "jcdvd" = "circ_jcdvd",
                           "jccd" = "circ_jccd",
                           "accd" = "circ_accd",
                           "otherType" = "circ_otherType",
                           "nameOfDay" = "day_of_week",
                           "totHours" = "open_total_hours",
                           "totBranchOpen" = "open_total_branches",
                           "pubHealthClosure" = "closed_public_health",
                           "closedNotes" = "closed_notes",
                           "totconnect" = "wifi_total_connections",
                           "totMBused" = "wifi_total_MB",
                           "totTraffic" = "traffic_total",
                           "totAttnd" = "programs_monthly_attendance",
                           "totProg" = "program_monthly_total",
                           "totPCsessions" = "pc_total_sessions",
                           "totOverDRiveCko" = "circ_overdrive_cko",
                           "totOverDRiveBorwCnt" = "circ_overdrive_bwrct",
                           "peopleTestedStl" = "covid_test_tested_stl", 
                           "positivesStl" = "covid_test_positive_stl", 
                           "confirmedCasesStl" = "covid_test_confirmed_stl", 
                           "hospitalizationsStl" = "covid_test_hospital_stl", 
                           "deathsStl" = "covid_test_deaths_stl", 
                           "peopleTestedCnty" = "covid_test_tested_cnty", 
                           "positivesCnty" = "covid_test_positive_cnty", 
                           "confirmedCasesCnty" = "covid_test_confirmed_cnty", 
                           "hospitalizationsCnty" = "covid_test_hospital_cnty", 
                           "deathsCnty" = "covid_test_deaths_cnty", 
                           "mRNA_1" = "covid_vax_mRNA_1_cnty",
                           "mRNA_2" = "covid_vax_mRNA_2_cnty",
                           "booster" = "covid_vax_booster_cnty",
                           "JJdose" = "covid_vax_JJdose_cnty",
                           "Population" = "covid_vax_population_cnty",
                           "pandemicDayNumber" = "pandemic_day_number",
                           "pandemicYes" = "pandemic_yes",
                           "stayAtHome" = "pandemic_stay_at_home",
                           )

# remove unneeded columns

data %<>% select(-c(allTestResultsStl, probableCasesStl, allTestResultsCnty, probableCasesCnty))

#, program_monthly_total, programs_monthly_attendance

```

```{r}
#add unique month/year/month name columns

for(x in 1:nrow(data)){
  data$month[x] <- paste0(str_split(data$date[x], pattern = "-")[[1]][2])
  data$year[x] <- paste0(str_split(data$date[x], pattern = "-")[[1]][1])
  data$month_name[x] <- paste0(str_split(data$date[x], pattern = "-")[[1]][2])
}

data$month_name %<>% recode("01" = "January",
                            "02" = "February",
                            "03" = "March",
                            "04" = "April",
                            "05" = "May",
                            "06" = "June",
                            "07" = "July",
                            "08" = "August",
                            "09" = "September",
                            "10" = "October",
                            "11" = "November",
                            "12" = "December")


data$pc_avg_sessions_hour <- round((data$pc_total_sessions / data$open_total_hours), 2)

# select and order columns


data %<>% select(c(date, circ_total_physical, circ_acbk, circ_jcbk, circ_accd, circ_pkbknh, circ_acdvd, circ_acdvd, circ_jccd,
                   circ_otherType, circ_overdrive_cko, circ_overdrive_bwrct, wifi_total_connections,wifi_total_MB, traffic_total,
                   pc_total_sessions, pc_avg_sessions_hour, program_monthly_total, programs_monthly_attendance,
                   year, month, month_name, day_of_week, open_total_hours, open_total_branches, closed_public_health, closed_notes,
                   covid_test_tested_stl, covid_test_positive_stl, covid_test_confirmed_stl, covid_test_hospital_stl, covid_test_deaths_stl, 
                   covid_test_tested_cnty, covid_test_positive_cnty, covid_test_confirmed_cnty, covid_test_hospital_cnty, covid_test_deaths_cnty, 
                   covid_vax_mRNA_1_cnty, covid_vax_mRNA_2_cnty, covid_vax_JJdose_cnty, covid_vax_booster_cnty, covid_vax_population_cnty, pandemic_yes, pandemic_day_number, 
                   pandemic_stay_at_home))

data[is.na(data)] <- 0

```


# Save daily data set

```{r}

dataDaily <- data
```

# write daily and monthly .csv files

```{r}
write.table(x = dataDaily, file = paste0(outFolder, "SPL_Full_Data_Daily_Updated_Again_18-21.csv"), quote = FALSE, sep = ",",
            row.names = FALSE, col.names = TRUE, na = "")
```














